The role of aneuploidy in the evolution of cancer drug resistance

### 2023-06-21 EvolutionaryRescue1.tex
moved survival probability calculations to Methods

\newcommand{\daniel}[1]{\textcolor{blue}{#1}}

@@

\daniel{I think this section has too much of the mathematical derivations, which should instead be in something like Methods. It should focus on just the results.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Survival probability}

To analyze evolutionary rescue in this model, we use the framework of \emph{multitype branching processes} \citep{harris1963theory}. 
This allows us to find explicit expressions for the \emph{survival probability}: the probability that a lineage descended from a single cell does not become extinct.

Let $p_w$, $p_a$, and $p_m$ be the survival probabilities of a population consisting initially of single wildtype cell, aneuploid cell, or mutant cell, respectively.
The complements $1-p_w$, $1-p_a$, and $1-p_m$ are the extinction probabilities, which satisfy each its respective equation,
\begin{equation} \label{extinction_prob}
\begin{aligned}
1-p_w = &\frac{\mu_w}{\lambda_w+\mu_w+u+v} + 
		  \frac{u}{\lambda_w+\mu_w+u+v}\left(1-p_a\right) + \\
		  & \frac{\lambda_w}{\lambda_w+\mu_w+u+v}\left(1-p_w\right)^2 +
		  \frac{v}{\lambda_w+\mu_w+u+v}\left(1-p_m\right) ,\\
1-p_a = &\frac{\mu_a}{\lambda_a+\mu_a+v}+\frac{v}{\lambda_a+\mu_a+v}\left(1-p_m\right)+\frac{\lambda_a}{\lambda_a+\mu_a+v}\left(1-p_a\right)^2 ,\\
1-p_m = &\frac{\mu_m}{\lambda_m+\mu_m}+\frac{\lambda_m}{\lambda_m+\mu_m}\left(1-p_m\right)^2 .	 
\end{aligned}
\end{equation}

@@

>>>>

\newcommand{\daniel}[1]{\textcolor{blue}{#1}}
\newcommand{\presc}{p_\text{rescue}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\section{Methods}

\subsection*{Calculation of survival probabilities of single lineages}

Here we calculate the \emph{survival probability}, the probability that a lineage descended from a single cell does not become extinct.
Let $p_w$, $p_a$, and $p_m$ be the survival probabilities of a lineage started by a single wildtype cell, aneuploid cell, or mutant cell, respectively.
We will neglect density dependence, allowing us to treat the lineages as \emph{multitype branching processes} \citep{harris1963theory}.
The extinction probabilities $1-p_w$, $1-p_a$, and $1-p_m$ then satisfy the equations:
\begin{equation} \label{extinction_prob}
\begin{aligned}
1-p_w = &\frac{\mu_w}{\lambda_w+\mu_w+u+v} + 
		  \frac{u}{\lambda_w+\mu_w+u+v}\left(1-p_a\right) + \\
		  & \frac{\lambda_w}{\lambda_w+\mu_w+u+v}\left(1-p_w\right)^2 +
		  \frac{v}{\lambda_w+\mu_w+u+v}\left(1-p_m\right) ,\\
1-p_a = &\frac{\mu_a}{\lambda_a+\mu_a+v}+\frac{v}{\lambda_a+\mu_a+v}\left(1-p_m\right)+\frac{\lambda_a}{\lambda_a+\mu_a+v}\left(1-p_a\right)^2 ,\\
1-p_m = &\frac{\mu_m}{\lambda_m+\mu_m}+\frac{\lambda_m}{\lambda_m+\mu_m}\left(1-p_m\right)^2 .	 
\end{aligned}
\end{equation}
